---
title: "STATS 101C Classification Report"
author: "Eric Chu, Van-An Lam, Ava Adlao, Liam Martin-McClay, Pavan Sah"
output:
  pdf_document: default
---


# Set-Up
If you want to knit the PDF but don't want to run the code chunks, set eval = FALSE.
```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)

set.seed(161)
```

## Import Libraries and Data
```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(readr)

train <- read_csv("../Data/train_class.csv")
test <- read_csv("../Data/test_class.csv")
```

# Intro

Many models are created to attempt to predict presidential election winners. The purpose of this model is to predict the winning presidential candidate per county using the 2020 election as the training and testing data. All data was acquired from the US Census Bureau.

We believe a county's type (i.e. urban or rural) is its most substantial predictor, as urban counties tend to exclusively vote Democrat. Furthermore, we also believe the variables associated with the county's result is the age and education level, as younger and more educated voters tend to vote Democrat, with vice versa demographics typically voting Republican.

# EDA

# Preproessing

# Recipes

# Models

# Final Model

# Appendix

## Final Model Script

## Team Contributions